OneD: increasing reproducibility of Hi-C Samples with abnormal karyotypes

### 2018-01-14 manuscript_nar.tex
More changes due to R2

The results so far suggest that the copy number correction lowers the
performance of \textit{OneD}. A possible limitation in Hi-C is the absence
of ground truth. We thus used simulations to generate matrices with
defined karyotypic aberrations (see Materials and Methods) and tested the
capacity of the correction methods to distinguish different cell types
(Supplementary Figure~12). On this test, \textit{HiCNorm} and \textit{OneD}
performed best. In comparison, \textit{OneD+CN} performed less well,
suggesting that correcting the copy number attenuates the biological
signal. This is corroborated by the fact that the copy number correction
alone has a lower performance than no correction at all. Thus, copy number
correction can indeed have side effects that lower the biological signal.

@@

We reasoned that \textit{OneD} should handle such cases gracefully,
because the signal is formally equivalent to an aberrant karyotype where
most of the genome is present in zero copy. We thus evaluated the
performance of \textit{OneD} on a Capture Hi-C data set at 5~Kbp
resolution centered on a 6~Mbp domain of chromosome 6 in T47D cells. We
found that the reproducibility score between replicates was higher after
normalization with \textit{OneD} than with matrix-balancing methods
(figure~\ref{fig:chic}A and Supplementary Figure~13). The performance
of \textit{OneD} was only slightly lower than the raw data, suggesting
that it barely disturbs the biological signal. Other tools were not
included beause they had prohibitive run time.

@@

To better understand these results, we also investigated the influence of
bin size and sequencing depth on the running time. We performed a
benchmark on the sample with highest sequencing depth in the data set
(HIC070, see Supplementary Table~1) at different bin sizes while
downsampling the contact matrices to mimic lower depths. On this
test, we included only the methods competing with \textit{OneD} in terms
of speed, \textit{i.e.} the matrix-balancing methods.
Figure~\ref{fig:times}B shows that the running time of \textit{OneD} does
not depend on the sequencing depth. Strikingly, \textit{OneD} also gives
nearly identical results at very different sequencing depths
(Supplementary Figure~14). \textit{vanilla} was the overall fastest method
and \textit{OneD} was usually faster than \textit{ICE} and \textit{KR}
at high but not at low sequencing depth. At low resolution (100~Kbp) the
speed advantage of \textit{OneD} appeared from low sequencing depth, but
at high resolution resolution (5~Kbp) it appeared only at high sequencing
depth.

>>>>

The results so far suggest that the copy number correction lowers the
performance of \textit{OneD}. A possible limitation in Hi-C is the absence
of ground truth. We thus used simulations to generate matrices with
defined karyotypic aberrations (see Materials and Methods) and tested the
capacity of the correction methods to distinguish different cell types
(Supplementary Figures~12 and 13). On this test, \textit{HiCNorm} and \textit{OneD}
performed best. In comparison, \textit{OneD+CN} performed less well,
suggesting that correcting the copy number attenuates the biological
signal. This is corroborated by the fact that the copy number correction
alone has a lower performance than no correction at all. Thus, copy number
correction can indeed have side effects that lower the biological signal.

@@

We reasoned that \textit{OneD} should handle such cases gracefully,
because the signal is formally equivalent to an aberrant karyotype where
most of the genome is present in zero copy. We thus evaluated the
performance of \textit{OneD} on a Capture Hi-C data set at 5~Kbp
resolution centered on a 6~Mbp domain of chromosome 6 in T47D cells. We
found that the reproducibility score between replicates was higher after
normalization with \textit{OneD} than with matrix-balancing methods
(figure~\ref{fig:chic}A and Supplementary Figure~14). The performance
of \textit{OneD} was only slightly lower than the raw data, suggesting
that it barely disturbs the biological signal. Other tools were not
included beause they had prohibitive run time.

@@

To better understand these results, we also investigated the influence of
bin size and sequencing depth on the running time. We performed a
benchmark on the sample with highest sequencing depth in the data set
(HIC070, see Supplementary Table~1) at different bin sizes while
downsampling the contact matrices to mimic lower depths. On this
test, we included only the methods competing with \textit{OneD} in terms
of speed, \textit{i.e.} the matrix-balancing methods.
Figure~\ref{fig:times}B shows that the running time of \textit{OneD} does
not depend on the sequencing depth. Strikingly, \textit{OneD} also gives
nearly identical results at very different sequencing depths
(Supplementary Figure~15). \textit{vanilla} was the overall fastest method
and \textit{OneD} was usually faster than \textit{ICE} and \textit{KR}
at high but not at low sequencing depth. At low resolution (100~Kbp) the
speed advantage of \textit{OneD} appeared from low sequencing depth, but
at high resolution resolution (5~Kbp) it appeared only at high sequencing
depth.
